### 2012-09-26 doc/prelim/document/chapters/nonlinear_problem/nonlinear_problem.tex
finished draft on nonlinear section

\subsection{Residual and Jacobian Generation}
\label{subsec:fanm_generation}

\subsection{Jacobian Storage vs. Subspace Storage and Restarts}
\label{subsec:fanm_storage}
To gauge the memory benefits of a FANM method over a Krylov-based
scheme, we must look at the storage requirements of the Jacobian
matrix as compared to the Krylov subspace vectors for sparse linear
problems. Saad's text provides us relations for both of these cases
\citep{saad_iterative_2003}. Beginning with the Jacobian storage, for
sparse problems production linear algebra library implementations
utilize \textit{compressed row storage} to efficiently store the
non-zero components of a sparse matrix while still allowing for all
standard parallel matrix computations to be performed. Per
\S~\ref{sec:parallel_krylov_methods}, recall that efficient parallel
matrix-vector multiplications rely on processors knowing which other
processors contain their neighboring matrix and vector
elements. Typically this information is represented by a graph which
must be stored along with the matrix elements themselves. To store the
elements, consider the following sparse matrix:
\[
  \ve{A} =
  \begin{bmatrix}
    2 & 0 & 8 & 0 & 0 & 0 \\ 
    4 & 5 & 0 & 1 & 0 & 0 \\ 
    0 & 2 & 1 & 0 & 1 & 0 \\ 
    0 & 0 & 3 & 7 & 0 & 2 \\ 
    0 & 0 & 0 & 4 & 9 & 0 \\
    0 & 0 & 0 & 0 & 9 & 1 \\
  \end{bmatrix}\:.
\]
The compressed form of \ve{A} would then be:
\[
  \begin{array}{lccccccccccccccc}
    \ve{A}\ values: & 2 & 8 & 4 & 5 & 1 & 2 & 1 & 1 & 3 & 7 & 2 & 4 & 9
    & 9 & 1 \\ column: & 1 & 3 & 1 & 2 & 4 & 2 & 3 & 5 & 3 & 4 & 6 & 4
    & 5 & 5 & 6 \\ row\ start: & 3 & 5 & 9 & 12 & 14 & 16 \\
  \end{array}
\]
where the first row contains the non-zero values of the matrix
$\ve{A}$ starts at, the second row contains the column index of those
values, and the third row contains the index of the A value and column
index rows that each matrix row starts at. For this simple example, we
actually break even by storing 36 elements in the full matrix and 36
pieces of data in the compressed format. However, for large, sparse
matrices there is a significant storage savings using the compressed
format.

\subsection{Parallel FANM Implementation}
\label{subsec:parallel_fanm}

>>>>

\subsection{Jacobian Storage vs. Subspace Storage and Restarts}
\label{subsec:fanm_storage}
To gauge the memory benefits of a FANM method over a Krylov-based
scheme, we must look at the storage requirements of the Jacobian
matrix as compared to the Krylov subspace vectors for sparse linear
problems. Saad's text provides us relations for both of these cases
\citep{saad_iterative_2003}. Beginning with the Jacobian storage, for
sparse problems production linear algebra library implementations
utilize \textit{compressed row storage} to efficiently store the
non-zero components of a sparse matrix while still allowing for all
standard parallel matrix computations to be performed. Per
\S~\ref{sec:parallel_krylov_methods}, recall that efficient parallel
matrix-vector multiplications rely on processors knowing which other
processors contain their neighboring matrix and vector
elements. Typically this information is represented by a graph which
must be stored along with the matrix elements themselves. To store the
elements, consider the following sparse matrix:
\[
  \ve{A} =
  \begin{bmatrix}
    2 & 0 & 8 & 0 & 0 & 0 \\ 
    4 & 5 & 0 & 1 & 0 & 0 \\ 
    0 & 2 & 1 & 0 & 1 & 0 \\ 
    0 & 0 & 3 & 7 & 0 & 2 \\ 
    0 & 0 & 0 & 4 & 9 & 0 \\
    0 & 0 & 0 & 0 & 9 & 1 \\
  \end{bmatrix}\:.
\]
The compressed form of \ve{A} would then be:
\[
  \begin{array}{lccccccccccccccc}
    \ve{A}\ values: & 2 & 8 & 4 & 5 & 1 & 2 & 1 & 1 & 3 & 7 & 2 & 4 & 9
    & 9 & 1 \\ column: & 1 & 3 & 1 & 2 & 4 & 2 & 3 & 5 & 3 & 4 & 6 & 4
    & 5 & 5 & 6 \\ row\ start: & 1 & 3 & 6 & 9 & 12 & 14 & 16 \\
  \end{array}
\]
where the first row contains the non-zero values of the matrix
$\ve{A}$ starts at, the second row contains the column index of those
values, and the third row contains the index of the A value and column
index rows that each matrix row starts at ending with the start of the
next row. For this simple example, we actually break even by storing
36 elements in the full matrix and 36 pieces of data in the compressed
format. However, for large, sparse matrices there is a significant
storage savings using the compressed format.

In his discussion on orthogonalization schemes for Krylov subspaces,
he provides us with space and time complexities for these
operations. To find a solution vector $\ve{x} \in \mathbb{R}^{N \times
  N}$ with $m$ Krylov iterations (and therefore $m+1$ subspace
vectors) the storage requirement is then $(m+1) N$ for most common
orthogonalization techniques. Using our simple example above (even
though we did not see any gains in storage over the full matrix
representation), if it takes 10 GMRES iterations using Arnoldi
orthogonalization to converge to a solution, then a total of 66 pieces
of data are required to be stored instead of 36 for the explicitly
formed matrix case. For larger sparse matrices, this disparity in
memory requirements will be even greater, especially for
ill-conditioned systems that require many GMRES iterations to
converge.

\subsection{Parallel FANM Implementation}
\label{subsec:parallel_fanm}
Like the parallelization of the MCSA method described in
\S~\ref{sec:parallel_stochastic_methods}, a parallel FANM method
relies on a basic set of parallel matrix-vector operations as well as
the global residual and Jacobian assembly procedure described in
\S~\ref{subsubsec:automatic_differentiation}. Consider the Newton
iteration scheme in Eq~(\ref{eq:newton_iteration}). We must first
assemble the linear system in parallel through the element-wise
function evaluations to generate both the global Jacobian operator and
the global residual vector on the right hand side. Per Bartlett's
work, efficient and automated parallel mechanisms are available to do
this through a sequence of scatter/gather operations. With these tools
available for residual and Jacobian generation, the remainder of the
parallel procedure is simple, with the linear Newton correction system
solved using the parallel MCSA method as previously described and the
Newton correction applied to the previous iterate's solution through a
vector update operation.

As Newton methods are formulated independent of the inner linear
solver generating the Newton corrections, the actual performance of
the nonlinear iterations using MCSA is expected to be similar to that
of traditional Newton-Krylov methods. Furthermore, we expect to
achieve numerically identical answers with a Newton-MCSA method as
other Newton methods and we should indeed verify this. The parallel
performance of such a method will inherently be bound to the parallel
Monte Carlo implementation of the linear solver as the parallel
operations at the nonlinear iteration level are identical to those
that you would perform with a Newton-Krylov method. Matrix-free
formulation will have different parallel performance than these
methods, and therefore we should compare FANM perfomance to JFNK-based
schemes. More important than performance in the situation is the
reduced memory pressure that a FANM implementation provides, as
discussed in \S~\ref{subsec:fanm_storage}, because a FANM method will
not generate a subspace in the linear solver and compressed storage
for sparse matrices are utilized, we expect significant memory savings
over Newton-Krylov methods. We must measure the memory utilization of
both of these methods in order to quantify their differences and
provide additional analysis of the FANM method's merits, or lack
thereof.
